---
output: html_document
---
<div style="display: flex; align-items: center; justify-content: flex-start;">
   <!-- Title on the left, big and bolded -->
   <h1 style="font-size: 35px; font-weight: bold; line-height: 150px; margin: 0;">
     Diabetes Overview and Insights
   </h1>
</div>

## Prevalence and Contributing Factors 
```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=NA, fig.width=12, fig.height=12}

#setwd("~/Desktop/web2/HealthDisparites")
# Load the necessary libraries
library(leaflet)
library(htmltools)
library(dplyr)
library(ggplot2)
library(sf)
library(stringr)
library(rnaturalearth)

# Get the map of Saudi Arabia using 'rnaturalearth' package
setwd("~/Desktop/a/nojoudalghamdi")
map <- ne_states(country = "Saudi Arabia", returnclass = "sf")

# Adjust region names to match data
map$name[1] = "Eastern Province"
map$name[2] = "Northern Borders"
map$name[3] = "Al Jawf"
map$name[4] = "Najran"
map$name[5] = "Asir"
map$name[6] = "Jazan"
map$name[7] = "Tabuk"
map$name[8] = "Madinah"
map$name[9] = "Makkah"
map$name[10] = "Riyadh"
map$name[11] = "Al Qassim"
map$name[12] = "Ha'il"
map$name[13] = "Al Bahah"

# Set the correct projection
map <- st_set_crs(map, 4326)

# Replace the data reading from Excel with the new data.frame
data <- data.frame(
  region = c("Riyadh", "Makkah", "Madinah", "Al Qassim", "Eastern Province", "Asir", "Tabuk", 
             "Ha'il", "Northern Borders", "Jazan", "Najran", "Al Bahah", "Al Jawf"),
  Diabetes = c(195, 194, 30, 24, 110, 57, 18, 8, 3, 27, 15, 40, 8),
  Abnormal_WHR = c(1666, 1505, 511, 292, 839, 550, 229, 153, 58, 283, 180, 105, 129),
  Dyslipdemia = c(227, 165, 26, 16, 87, 53, 9, 9, 4, 21, 12, 30, 9),
  Elevated_Glucose = c(345, 378, 63, 42, 96, 36, 75, 13, 1, 43, 14, 24, 3),
  Insufficient_Diet_and_Exercise = c(4193, 3904, 1162, 669, 1915, 1127, 455, 337, 141, 701, 344, 246, 264),
  High_BMI = c(1241, 992, 306, 212, 681, 380, 177, 82, 38, 171, 115, 102, 106)
)

# Ensure the data is ordered alphabetically by region name
data <- data[order(data$region), ]

# Merge the map data with the health data
map <- left_join(map, data, by = c("name" = "region"))

# Create a wrapped version of region names for better appearance
map$name_wrapped <- str_wrap(map$name, width = 10)

# Calculate centroids for label placement
map_centroids <- st_centroid(map)

# Define custom color palettes for each variable
palette_diabetes <- colorNumeric(
  palette = c("#f2f8fa", "#b4e0e2", "#88c4cd", "#73b2b5"),
  domain = map$Diabetes,
  na.color = "#D3D3D3"
)
palette_blood_glucose <- colorNumeric(
  palette = c("#f2f8fa", "#b4e0e2", "#88c4cd", "#73b2b5"),
  domain = map$Elevated_Glucose,
  na.color = "#D3D3D3"
)
palette_bmi <- colorNumeric(
  palette = c("#f2f8fa", "#b4e0e2", "#88c4cd", "#73b2b5"),
  domain = map$High_BMI,
  na.color = "#D3D3D3"
)
palette_diet_exercise <- colorNumeric(
  palette = c("#f2f8fa", "#b4e0e2", "#88c4cd", "#73b2b5"),
  domain = map$Insufficient_Diet_and_Exercise,
  na.color = "#D3D3D3"
)
palette_whr <- colorNumeric(
  palette = c("#f2f8fa", "#b4e0e2", "#88c4cd", "#73b2b5"),
  domain = map$Abnormal_WHR,
  na.color = "#D3D3D3"
)
palette_dyslipidemia <- colorNumeric(
  palette = c("#f2f8fa", "#b4e0e2", "#88c4cd", "#73b2b5"),
  domain = map$Dyslipdemia,
  na.color = "#D3D3D3"
)

# Function to create the interactive map
create_map <- function(variable, palette) { 
  leaflet(data = map) %>%
    addProviderTiles(providers$CartoDB.Positron) %>%
    addPolygons(
      fillColor = ~palette(variable),
      weight = 1,
      opacity = 1,
      color = "black",
      dashArray = "3",
      fillOpacity = 0.7,
      highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE),
      label = ~paste0("Cases: ", round(variable, 2))  # Only show the value as "Cases"
    )
}

# Create the maps with custom color palettes
map_diabetes <- create_map(map$Diabetes, palette_diabetes)
map_blood_glucose <- create_map(map$Elevated_Glucose, palette_blood_glucose)
map_bmi <- create_map(map$High_BMI, palette_bmi)
map_diet_exercise <- create_map(map$Insufficient_Diet_and_Exercise, palette_diet_exercise)
map_whr <- create_map(map$Abnormal_WHR, palette_whr)
map_dyslipidemia <- create_map(map$Dyslipdemia, palette_dyslipidemia)

# Combine the maps in a grid layout with titles centered above each map
htmltools::browsable(
  tagList(
    div(
      style = "display: grid; grid-template-columns: repeat(2, 1fr); grid-template-rows: repeat(3, 1fr); gap: 20px; justify-items: center; padding: 20px; max-width: 80%; box-sizing: border-box;",
      # Title and map for Diabetes
      div(
        style = "display: flex; flex-direction: column; align-items: center; width: 100%;",
        div(style = "color: #4f6d7a; font-size: 18px; font-weight: bold; text-align: center; margin-bottom: 10px;", "Diabetes"),
        div(style = "width: 100%; height: 100%;", map_diabetes)
      ),
      # Title and map for Elevated Blood Glucose
      div(
        style = "display: flex; flex-direction: column; align-items: center; width: 100%;",
        div(style = "color: #4f6d7a; font-size: 18px; font-weight: bold; text-align: center; margin-bottom: 10px;", "Elevated Blood Glucose"),
        div(style = "width: 100%; height: 100%;", map_blood_glucose)
      ),
      # Title and map for High BMI
      div(
        style = "display: flex; flex-direction: column; align-items: center; width: 100%;",
        div(style = "color: #4f6d7a; font-size: 18px; font-weight: bold; text-align: center; margin-bottom: 10px;", "High BMI"),
        div(style = "width: 100%; height: 100%;", map_bmi)
      ),
      # Title and map for Insufficient Diet and Exercise
      div(
        style = "display: flex; flex-direction: column; align-items: center; width: 100%;",
        div(style = "color: #4f6d7a; font-size: 18px; font-weight: bold; text-align: center; margin-bottom: 10px;", "Insufficient Diet and Exercise"),
        div(style = "width: 100%; height: 100%;", map_diet_exercise)
      ),
      # Title and map for Abnormal WHR
      div(
        style = "display: flex; flex-direction: column; align-items: center; width: 100%;",
        div(style = "color: #4f6d7a; font-size: 18px; font-weight: bold; text-align: center; margin-bottom: 10px;", "Abnormal WHR"),
        div(style = "width: 100%; height: 100%;", map_whr)
      ),
      # Title and map for Dyslipidemia
      div(
        style = "display: flex; flex-direction: column; align-items: center; width: 100%;",
        div(style = "color: #4f6d7a; font-size: 18px; font-weight: bold; text-align: center; margin-bottom: 10px;", "Dyslipidemia"),
        div(style = "width: 100%; height: 100%;", map_dyslipidemia)
      )
    )
  )
)

# cd ~/Desktop/web/HealthDisparites
# git init
# git add .
# git commit -m "first"
# git push -u origin main

```
- High prevalence in **Makkah, Riyadh, Eastern Province** due to urbanization, sedentary lifestyles, and poor diet.
- **Madinah, Asir, Al Qassim** show elevated rates from poor diet, lack of exercise, and abnormal WHR.
- Regions like **Tabuk** and **Madinah** show high blood glucose, and **Asir** has high BMI.
<br>

## Summary Statistics and Demographic Trends
```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=NA, fig.width=12, fig.height=12}
library(gt)
table_data <- data.frame(
  Condition = "Diabetes",
  N = 13,
  Mean = 56.08,
  `Standard Deviation` = 67.467
)

# Render the table with gt
table_data %>%
  gt() %>%
  tab_header(title = "Descriptive Statistics") %>%
  tab_source_note(
    source_note = "Descriptive statistics of diabetes prevalence in Saudi Arabia, 2019."
  )
```
- Mean prevalence: 56.08%, with high variability (SD: 67.47%).
- 2030 projection: ~72.3 million cases in Saudi Arabia.

```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=NA, fig.width=12, fig.height=12}

library(plotly)
library(htmltools)

# Define the dataset for Diabetes cases by Age
diabetes_data <- data.frame(
  Age = c("15-29", "30-44", "45-59", "60-69", "70-79", "80+"),
  Diabetes = c(60, 165, 247, 166, 69, 21)
)

# Create bar chart with hover displaying only the number of cases and no title on the graph
diabetes_chart <- plot_ly(
  data = diabetes_data,
  x = ~Age,
  y = ~Diabetes,
  type = "bar",
  marker = list(color = "#b4e0e2"),
  hovertemplate = "Cases: %{y}<extra></extra>"  # Customize hover text
) %>% layout(
  title = "",  # Remove the title from the graph
  xaxis = list(
    title = list(text = "Age Group"),
    tickfont = list(size = 12, color = "#4f6d7a")
  ),
  yaxis = list(
    title = list(text = "Number of Cases"),
    tickfont = list(size = 12, color = "#4f6d7a")
  ),
  width = 400,  # Adjust the width
  height = 400  # Adjust the height of the bar chart
)

# Define the dataset for Gender and Diabetes cases
gender_diabetes_data <- data.frame(
  Gender = c("Male", "Female"),
  Diabetes = c(408, 321)
)

# Create pie chart for Gender and Diabetes cases
gender_diabetes_pie_chart <- plot_ly(
  data = gender_diabetes_data,
  labels = ~Gender,  # Labels for Male and Female
  values = ~Diabetes,  # Values for the number of cases
  type = "pie",
  marker = list(colors = c("#b4e0e2", "#88c4cd")),  # Same colors as the bar chart
  hovertemplate = "Cases: %{value}<extra></extra>",  # Show only the number of cases on hover
  textinfo = "label",  # Only show labels (Male and Female), no numbers inside the pie
  textposition = "inside"  # Position the text inside the pie chart slices
) %>% layout(
  title = "",  # No title on the graph
  showlegend = FALSE,  # Hide the legend since the labels are already inside the pie chart
  width = 400,  # Set the width of the pie chart
  height = 400  # Adjust the height of the pie chart
)

# Display both charts next to each other with titles above
htmltools::browsable(
  tagList(
    div(
      style = "display: flex; justify-content: space-between; height: 500px; padding-top: 20px;",  # Adjust padding to add space above
      # First chart (Bar chart for Diabetes Cases by Age)
      div(
        style = "width: 400px; height: 400px;",  # Set width and height for the first chart
        div(
          style = "font-weight: bold; text-align: center; margin: 0; padding: 10px 0; font-size: 17px;",  # Add padding to avoid overlap
          "Diabetes Cases by Age"  # Title for the bar chart
        ),
        diabetes_chart  # Bar chart for Diabetes by Age
      ),
      # Second chart (Pie chart for Diabetes Cases by Gender)
      div(
        style = "width: 400px; height: 400px;",  # Set width and height for the second chart
        div(
          style = "font-weight: bold; text-align: center; margin: 0; padding: 10px 0; font-size: 17px;",  # Add padding to avoid overlap
          "Diabetes Cases by Gender"  # Title for the pie chart
        ),
        gender_diabetes_pie_chart  # Pie chart for Diabetes by Gender
      )
    )
  )
)

```
- Highest prevalence in individuals aged 45–59, accounting for 33.9% of total cases.
- The second highest prevalence is in the age group of 60–69, constituting 22.8%.
- Gender distribution indicates a higher prevalence in males (55.97%) compared to females (44.03%), suggesting a significant gender disparity.
<br>

## Correlation Insights
```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=NA, fig.width=12, fig.height=12}
library(gt)

# Create the data frame
table_data <- data.frame(
  Factor = c("Abnormal WHR", "Elevated Blood Glucose", "Dyslipidemia", "High BMI", "Insufficient Diet and Exercise"),
  Pearson_Correlation = c("0.979**", "0.952**", "0.979**", "0.980**", "0.981**"),
  Sig_2_tailed = c("<0.001", "<0.001", "<0.001", "<0.001", "<0.001"),
  N = rep(13, 5)
)

# Render the table with gt
table_data %>%
  gt() %>%
  tab_header(
    title = "Pearson Correlation Coefficients Between Diabetes and Influencing Factors"
  ) %>%
  cols_label(
    Factor = "Factors",
    Pearson_Correlation = "Pearson Correlation",
    Sig_2_tailed = "Sig. (2-tailed)",
    N = "N"
  ) %>%
  tab_source_note(
    source_note = "Pearson correlation coefficients between diabetes and influencing factors."
  )
```
- Strong correlation with **dyslipidemia**, **BMI**, and **poor diet/exercise**.
- Weaker correlation with **elevated blood glucose** and **abnormal WHR**.
<br> 

## Regression Insights: Fixed vs Random Effects
```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=NA, fig.width=12, fig.height=12}
library(gt)

# Create the data frame
table_data <- data.frame(
  Variable = c("β₀ (Intercept)", "β₁ (Abnormal WHR)", "β₂ (Elevated Blood Glucose)", 
               "β₃ (Dyslipidemia)", "β₄ (High BMI)", "β₅ (Insufficient Diet and Exercise)"),
  Mean = c(0.046, 0.002, 0.008, 0.008, 0.002, -0.003),
  Standard_Deviation = c(0.207, 0.007, 0.006, 0.011, 0.004, 0.003),
  Credible_Interval = c("(-0.381, 0.436)", "(-0.011, 0.016)", "(-0.004, 0.019)", 
                        "(-0.014, 0.030)", "(-0.006, 0.011)", "(-0.008, 0.002)")
)

# Render the table with gt
table_data %>%
  gt() %>%
  tab_header(
    title = "Bayesian Posterior Inference for the Fixed Effects in the Diabetes Model"
  ) %>%
  cols_label(
    Variable = "Fixed Effect",
    Mean = "Mean",
    Standard_Deviation = "Standard Deviation",
    Credible_Interval = "95% Credible Interval"
  ) %>%
  tab_source_note(
    source_note = "Bayesian posterior inference for the fixed effects in the diabetes model."
  )

library(gt)

# Create the data frame
table_data <- data.frame(
  Variable = c("τᵤ", "τᵥ"),
  Mean = c(9.92, 22036.44),
  Standard_Deviation = c(6.46, 24106.86),
  Credible_Interval = c("(2.44, 26.69)", "(1473.44, 86045.78)")
)

# Render the table with gt
table_data %>%
  gt() %>%
  tab_header(
    title = "Bayesian Posterior Inference for the Random Effects in the Diabetes Model"
  ) %>%
  cols_label(
    Variable = "Random Effect",
    Mean = "Mean",
    Standard_Deviation = "Standard Deviation",
    Credible_Interval = "95% Credible Interval"
  ) %>%
  tab_source_note(
    source_note = "Bayesian posterior inference for the random effects in the diabetes model."
  )

```
- Positive but non-significant association with **abnormal WHR**, **elevated blood glucose**, and **dyslipidemia**.
- Local factors more significant than geographic proximity in determining diabetes risk.
<br> 

## Risk Mapping and Statistical Insights
```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=NA, fig.width=12, fig.height=12}

# Load the necessary libraries
library(leaflet)
library(dplyr)
library(sf)
library(rnaturalearth)

# Define the data
map_data <- data.frame(
  name = c("Al Bahah", "Al Jawf", "Al Qassim", "Asir", "Eastern Province", 
           "Ha'il", "Jazan", "Madinah", "Makkah", "Najran", 
           "Northern Borders", "Riyadh", "Tabuk"),
  RR = c(3.9124206, 0.7054240, 0.7775904, 1.2440027, 0.9586025, 
         0.6217034, 0.8789209, 0.6136727, 1.0130544, 1.1888723, 
         0.6522526, 1.0257210, 1.0190876),
  EP = c(0.9999995812, 0.0363428253, 0.0230961909, 0.8107168262, 0.0680944281, 
         0.0045105158, 0.0875673861, 0.0002221726, 0.1182955821, 0.5934219066, 
         0.0296411917, 0.1539461439, 0.3256815848),
  structured_effect = c(1.042330155, -0.378137622, -0.159831323, 0.274990220, 0.001125799, 
                        -0.371859912, 0.124295371, -0.047432847, 0.350303145, 0.121362623, 
                        -0.413673421, -0.109093635, -0.434268320),
  unstructured_effect = c(6.151902e-04, 1.111757e-05, 1.198330e-05, -4.594934e-05, 1.872479e-04, 
                          -2.781093e-04, -1.639213e-04, 1.807496e-04, 2.122470e-04, 
                          6.445639e-05, -3.003219e-04, -2.820046e-04, -2.115819e-04)
)

# Get the map of Saudi Arabia using the 'rnaturalearth' package
map <- ne_states(country = "Saudi Arabia", returnclass = "sf")

# Adjust region names to match data
map$name[1] = "Eastern Province"
map$name[2] = "Northern Borders"
map$name[3] = "Al Jawf"
map$name[4] = "Najran"
map$name[5] = "Asir"
map$name[6] = "Jazan"
map$name[7] = "Tabuk"
map$name[8] = "Madinah"
map$name[9] = "Makkah"
map$name[10] = "Riyadh"
map$name[11] = "Al Qassim"
map$name[12] = "Ha'il"
map$name[13] = "Al Bahah"

# Set the correct projection
map <- st_set_crs(map, 4326)

# Merge the map data with the new data
map <- left_join(map, map_data, by = c("name" = "name"))

# Create custom color palettes for each variable
palette_rr <- colorNumeric(
  palette = c("#f2f8fa", "#b4e0e2", "#88c4cd", "#73b2b5"),
  domain = map$RR,
  na.color = "#D3D3D3"
)

palette_ep <- colorNumeric(
  palette = c("#f2f8fa", "#b4e0e2", "#88c4cd", "#73b2b5"),
  domain = map$EP,
  na.color = "#D3D3D3"
)

palette_structured_effect <- colorNumeric(
  palette = c("#f2f8fa", "#b4e0e2", "#88c4cd", "#73b2b5"),
  domain = map$structured_effect,
  na.color = "#D3D3D3"
)

palette_unstructured_effect <- colorNumeric(
  palette = c("#f2f8fa", "#b4e0e2", "#88c4cd", "#73b2b5"),
  domain = map$unstructured_effect,
  na.color = "#D3D3D3"
)

# Function to create the interactive map
# Function to create the interactive map without the legend
create_map_no_legend <- function(variable, palette, title) {
  leaflet(data = map) %>%
    addProviderTiles(providers$CartoDB.Positron) %>%
    addPolygons(
      fillColor = ~palette(variable),
      weight = 1,
      opacity = 1,
      color = "black",
      dashArray = "3",
      fillOpacity = 0.7,
      highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE),
      label = ~paste0(title, ": ", round(variable, 2))
    )
}

# Create maps for each variable without the legend
map_rr_no_legend <- create_map_no_legend(map$RR, palette_rr, "RR")
map_ep_no_legend <- create_map_no_legend(map$EP, palette_ep, "EP")
map_structured_effect_no_legend <- create_map_no_legend(map$structured_effect, palette_structured_effect, "Structured Effect")
map_unstructured_effect_no_legend <- create_map_no_legend(map$unstructured_effect, palette_unstructured_effect, "Unstructured Effect")

# Combine the maps in a grid layout with titles centered above each map
htmltools::browsable(
  tagList(
    div(
      style = "display: grid; grid-template-columns: repeat(2, 1fr); grid-template-rows: repeat(2, 1fr); gap: 20px; justify-items: center; padding: 20px; max-width: 80%; box-sizing: border-box;",
      # Title and map for RR
      div(
        style = "display: flex; flex-direction: column; align-items: center; width: 100%;",
        div(style = "color: #4f6d7a; font-size: 18px; font-weight: bold; text-align: center; margin-bottom: 10px;", "RR"),
        div(style = "width: 100%; height: 100%;", map_rr_no_legend)
      ),
      # Title and map for EP
      div(
        style = "display: flex; flex-direction: column; align-items: center; width: 100%;",
        div(style = "color: #4f6d7a; font-size: 18px; font-weight: bold; text-align: center; margin-bottom: 10px;", "EP"),
        div(style = "width: 100%; height: 100%;", map_ep_no_legend)
      ),
      # Title and map for Structured Effect
      div(
        style = "display: flex; flex-direction: column; align-items: center; width: 100%;",
        div(style = "color: #4f6d7a; font-size: 18px; font-weight: bold; text-align: center; margin-bottom: 10px;", "Structured Effect"),
        div(style = "width: 100%; height: 100%;", map_structured_effect_no_legend)
      ),
      # Title and map for Unstructured Effect
      div(
        style = "display: flex; flex-direction: column; align-items: center; width: 100%;",
        div(style = "color: #4f6d7a; font-size: 18px; font-weight: bold; text-align: center; margin-bottom: 10px;", "Unstructured Effect"),
        div(style = "width: 100%; height: 100%;", map_unstructured_effect_no_legend)
      )
    )
  )
)


# cd ~/Desktop/a/nojoudalghamdi
# git init
# git add .
# git commit -m "second"
# git push -u origin main

```
- **Al Bahah** shows high risk, while **Northern Borders** and **Riyadh** show lower risk.
- **Ep**: Regions like **Makkah**, **Eastern Province**, and **Riyadh** may show clustering of high-risk areas.
- **Structured random effects** show minimal spatial influence; **unstructured random effects** suggest local factors (e.g., lifestyle, diet) are more important.
