---
output: html_document
---
<div style="display: flex; align-items: center; justify-content: flex-start;">
   <!-- Title on the left, big and bolded -->
   <h1 style="font-size: 35px; font-weight: bold; line-height: 150px; margin: 0;">
     Stroke Overview and Insights
   </h1>
</div>

## Prevalence and Contributing Factors 
```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=NA, fig.width=12, fig.height=12}
# Load the necessary libraries
library(leaflet)
library(htmltools)
library(dplyr)
library(ggplot2)
library(sf)
library(stringr)
library(rnaturalearth)

# Get the map of Saudi Arabia using 'rnaturalearth' package
setwd("~/Desktop/web2/HealthDisparites")
map <- ne_states(country = "Saudi Arabia", returnclass = "sf")

# Adjust region names to match data
map$name[1] = "Eastern Province"
map$name[2] = "Northern Borders"
map$name[3] = "Al Jawf"
map$name[4] = "Najran"
map$name[5] = "Asir"
map$name[6] = "Jazan"
map$name[7] = "Tabuk"
map$name[8] = "Madinah"
map$name[9] = "Makkah"
map$name[10] = "Riyadh"
map$name[11] = "Al Qassim"
map$name[12] = "Ha'il"
map$name[13] = "Al Bahah"

# Set the correct projection
map <- st_set_crs(map, 4326)

# Replace the data reading from Excel with the new data.frame
data <- data.frame(
  region = c("Riyadh", "Makkah", "Madinah", "Al Qassim", "Eastern Province", "Asir", "Tabuk", 
             "Ha'il", "Northern Borders", "Jazan", "Najran", "Al Bahah", "Al Jawf"),
 Stroke = c(12, 11, 1, 0, 6, 0, 0, 0, 0, 0, 0, 0, 0),
  Total_Tobacco_Consumption = c(392, 273, 58, 14, 94, 60, 36, 23, 8, 34, 7, 15, 26),
  WPM2.5 = c(49.00536, 51.86058, 48.01454, 44.14183, 47.32641, 
             39.71297, 26.44076, 34.035, 31.05663, 45.1512, 
             54.90682, 40.56398, 25.87969),
  Dyslipdemia = c(227, 165, 26, 16, 87, 53, 9, 9, 4, 21, 12, 30, 9)
)

# Ensure the data is ordered alphabetically by region name
data <- data[order(data$region), ]

# Merge the map data with the health data
map <- left_join(map, data, by = c("name" = "region"))

# Create a wrapped version of region names for better appearance
map$name_wrapped <- str_wrap(map$name, width = 10)

# Calculate centroids for label placement
map_centroids <- st_centroid(map)

# Define custom color palettes for each variable
palette_Stroke <- colorNumeric(
  palette = c("#f2f8fa", "#b4e0e2", "#88c4cd", "#73b2b5"),
  domain = map$Hypertension,
  na.color = "#D3D3D3"
)
palette_Total_Tobacco_Consumption <- colorNumeric(
  palette = c("#f2f8fa", "#b4e0e2", "#88c4cd", "#73b2b5"),
  domain = map$Total_Tobacco_Consumption,
  na.color = "#D3D3D3"
)
palette_Dyslipdemia <- colorNumeric(
  palette = c("#f2f8fa", "#b4e0e2", "#88c4cd", "#73b2b5"),
  domain = map$Dyslipdemia,
  na.color = "#D3D3D3"
)
palette_WPM2.5 <- colorNumeric(
  palette = c("#f2f8fa", "#b4e0e2", "#88c4cd", "#73b2b5"),
  domain = map$WPM2.5,
  na.color = "#D3D3D3"
)

# Function to create the interactive map
create_map <- function(variable, palette) { 
  leaflet(data = map) %>%
    addProviderTiles(providers$CartoDB.Positron) %>%
    addPolygons(
      fillColor = ~palette(variable),
      weight = 1,
      opacity = 1,
      color = "black",
      dashArray = "3",
      fillOpacity = 0.7,
      highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE),
      label = ~paste0("Cases: ", round(variable, 2))  # Only show the value as "Cases"
    )
}

# Create the maps with custom color palettes
map_Stroke <- create_map(map$Stroke, palette_Stroke)
map_Total_Tobacco_Consumption <- create_map(map$Total_Tobacco_Consumption, palette_Total_Tobacco_Consumption)
map_Dyslipdemia <- create_map(map$Dyslipdemia, palette_Dyslipdemia)
map_WPM2.5 <- create_map(map$WPM2.5, palette_WPM2.5)

# Combine the maps in a grid layout with titles centered above each map
htmltools::browsable(
  tagList(
    div(
      style = "display: grid; grid-template-columns: repeat(2, 1fr); grid-template-rows: auto auto auto; gap: 20px; justify-items: center; padding: 20px; max-width: 80%; box-sizing: border-box;",
      
      # First row with 2 maps
      div(
        style = "display: flex; flex-direction: column; align-items: center; width: 100%;",
        div(style = "color: #4f6d7a; font-size: 18px; font-weight: bold; text-align: center; margin-bottom: 10px;", "Stroke"),
        div(style = "width: 100%; height: 100%;", map_Hypertension)
      ),
      div(
        style = "display: flex; flex-direction: column; align-items: center; width: 100%;",
        div(style = "color: #4f6d7a; font-size: 18px; font-weight: bold; text-align: center; margin-bottom: 10px;", "Total Tobacco Consumption"),
        div(style = "width: 100%; height: 100%;", map_Total_Tobacco_Consumption)
      ),
      
      # Second row with 2 maps
      div(
        style = "display: flex; flex-direction: column; align-items: center; width: 100%;",
        div(style = "color: #4f6d7a; font-size: 18px; font-weight: bold; text-align: center; margin-bottom: 10px;", "Dyslipdemia"),
        div(style = "width: 100%; height: 100%;", map_Dyslipdemia)
      ),
      # Third row with WPM2.5 centered (same size as above maps)
      div(
        style = "display: flex; flex-direction: column; align-items: center; width: 100%;",
        div(style = "color: #4f6d7a; font-size: 18px; font-weight: bold; text-align: center; margin-bottom: 10px;", "WPM2.5"),
        div(style = "width: 100%; height: 100%; max-width: 100%;", map_WPM2.5)
      )
    )
  )
)
```
- Higher stroke rates in **Makkah, Riyadh, Eastern Province** due to urbanization and pollution.
- **Northern regions** like **Al Jawf**, **Tabuk**, **Ha’il** have lower pollution.

## Summary Statistics and Demographic Trends
```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=NA, fig.width=12, fig.height=12}
library(gt)

# Create the data frame
table_data <- data.frame(
  Condition = "Stroke",
  N = 13,
  Mean = 2.31,
  Standard_Deviation = 4.404
)

# Render the table with gt
table_data %>%
  gt() %>%
  tab_header(
    title = "Descriptive Statistics of Stroke Prevalence in Saudi Arabia, 2019"
  ) %>%
  cols_label(
    Condition = "Condition",
    N = "N",
    Mean = "Mean",
    Standard_Deviation = "Standard Deviation"
  ) %>%
  tab_source_note(
    source_note = "Descriptive statistics of stroke prevalence in Saudi Arabia, 2019."
  )
```
- Mean prevalence: 2.31%, with variability (SD: 4.40%).
- 2030 projection: ~2.98 million cases.

```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=NA, fig.width=12, fig.height=12}
library(plotly)
library(htmltools)

# Define the dataset for Stroke cases by Age
stroke_data <- data.frame(
  Age = c("15-29", "30-44", "45-59", "60-69", "70-79", "80+"),
  Stroke = c(0, 4, 5, 11, 6, 4)
)

# Create bar chart for Stroke Cases by Age
stroke_age_chart <- plot_ly(
  data = stroke_data,
  x = ~Age,
  y = ~Stroke,
  type = "bar",
  marker = list(color = "#b4e0e2"),
  hovertemplate = "Cases: %{y}<extra></extra>"  # Customize hover text
) %>% layout(
  title = "",  # Remove the title from the graph
  xaxis = list(
    title = list(text = "Age Group"),
    tickfont = list(size = 12, color = "#4f6d7a")
  ),
  yaxis = list(
    title = list(text = "Number of Cases"),
    tickfont = list(size = 12, color = "#4f6d7a")
  ),
  width = 400,  # Adjust the width
  height = 400  # Adjust the height of the bar chart
)

# Define the dataset for Gender and Stroke cases
gender_stroke_data <- data.frame(
  Gender = c("Male", "Female"),
  Stroke = c(19, 8)
)

# Create pie chart for Gender and Stroke cases
gender_stroke_pie_chart <- plot_ly(
  data = gender_stroke_data,
  labels = ~Gender,  # Labels for Male and Female
  values = ~Stroke,  # Values for the number of cases
  type = "pie",
  marker = list(colors = c("#b4e0e2", "#88c4cd")),  # Same colors as the bar chart
  hovertemplate = "Cases: %{value}<extra></extra>",  # Show only the number of cases on hover
  textinfo = "label",  # Only show labels (Male and Female), no numbers inside the pie
  textposition = "inside"  # Position the text inside the pie chart slices
) %>% layout(
  title = "",  # No title on the graph
  showlegend = FALSE,  # Hide the legend since the labels are already inside the pie chart
  width = 400,  # Set the width of the pie chart
  height = 400  # Adjust the height of the pie chart
)

# Display both charts next to each other with titles above
htmltools::browsable(
  tagList(
    div(
      style = "display: flex; justify-content: space-between; height: 500px; padding-top: 20px;",  # Adjust padding to add space above
      # First chart (Bar chart for Stroke Cases by Age)
      div(
        style = "width: 400px; height: 400px;",  # Set width and height for the first chart
        div(
          style = "font-weight: bold; text-align: center; margin: 0; padding: 10px 0; font-size: 17px;",  # Add padding to avoid overlap
          "Stroke Cases by Age"  # Title for the bar chart
        ),
        stroke_age_chart  # Bar chart for Stroke by Age
      ),
      # Second chart (Pie chart for Stroke Cases by Gender)
      div(
        style = "width: 400px; height: 400px;",  # Set width and height for the second chart
        div(
          style = "font-weight: bold; text-align: center; margin: 0; padding: 10px 0; font-size: 17px;",  # Add padding to avoid overlap
          "Stroke Cases by Gender"  # Title for the pie chart
        ),
        gender_stroke_pie_chart  # Pie chart for Stroke by Gender
      )
    )
  )
)
```
- Highest prevalence in individuals aged 60–69, accounting for 36.7% of total cases.
- The second highest prevalence is in the age group of 45–59, constituting 16.7%.
- Gender distribution shows a higher prevalence among males (70.37%) compared to females (29.63%), indicating a more pronounced impact on the male population.

## Correlation Insights
```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=NA, fig.width=12, fig.height=12}
library(gt)

# Create the data frame
table_data <- data.frame(
  Factor = c("Total Tobacco Consumption", "WPM_{2.5}", "Dyslipidemia"),
  Pearson_Correlation = c("0.957**", "0.496", "0.969**"),
  Sig_2_tailed = c("<0.001", "0.085", "<0.001"),
  N = rep(13, 3)
)

# Render the table with gt
table_data %>%
  gt() %>%
  tab_header(
    title = "Pearson Correlation Coefficients Between Stroke and Influencing Factors"
  ) %>%
  cols_label(
    Factor = "Factors",
    Pearson_Correlation = "Pearson Correlation",
    Sig_2_tailed = "Sig. (2-tailed)",
    N = "N"
  ) %>%
  tab_source_note(
    source_note = "Pearson correlation coefficients between stroke and influencing factors."
  )
```
- Strong correlation with **dyslipidemia** and **total tobacco consumption**.
- Weaker correlation with **WPM$_{2.5}$** pollution.

## Regression Insights: Fixed vs Random Effects
```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=NA, fig.width=12, fig.height=12}
library(gt)

# Create the data frame
table_data <- data.frame(
  Variable = c("β₀ (Intercept)", "β₁ (Total Tobacco Consumption)", "β₂ (WPM_{2.5})", "β₃ (Dyslipidemia)"),
  Mean = c(-10.246, -0.032, 0.160, 0.067),
  Standard_Deviation = c(4.516, 0.015, 0.087, 0.030),
  Credible_Interval = c("(-19.102, -1.390)", "(-0.062, -0.002)", "(-0.010, 0.330)", "(0.008, 0.127)")
)

# Render the table with gt
table_data %>%
  gt() %>%
  tab_header(
    title = "Bayesian Posterior Inference for the Fixed Effects in the Stroke Model"
  ) %>%
  cols_label(
    Variable = "Fixed Effect",
    Mean = "Mean",
    Standard_Deviation = "Standard Deviation",
    Credible_Interval = "95% Credible Interval"
  ) %>%
  tab_source_note(
    source_note = "Bayesian posterior inference for the fixed effects in the stroke model."
  )

library(gt)

# Create the data frame
table_data <- data.frame(
  Variable = c("τᵤ", "τᵥ"),
  Mean = c(22041.23, 22042.08),
  Standard_Deviation = c(24160.08, 24163.00),
  Credible_Interval = c("(1471.37, 86188.70)", "(1471.39, 86197.14)")
)

# Render the table with gt
table_data %>%
  gt() %>%
  tab_header(
    title = "Bayesian Posterior Inference for the Random Effects in the Stroke Model"
  ) %>%
  cols_label(
    Variable = "Random Effect",
    Mean = "Mean",
    Standard_Deviation = "Standard Deviation",
    Credible_Interval = "95% Credible Interval"
  ) %>%
  tab_source_note(
    source_note = "Bayesian posterior inference for the random effects in the stroke model."
  )

```
- No significant relationship with **WPM$_{2.5}$**.
- Strong positive relationship with **dyslipidemia** and **tobacco consumption**.

## Risk Mapping and Statistical Insights
```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=NA, fig.width=12, fig.height=12}

# Load the necessary libraries
library(leaflet)
library(dplyr)
library(sf)
library(rnaturalearth)

# Define the data
map_data <- data.frame(
  name = c("Al Bahah", "Al Jawf", "Al Qassim", "Asir", "Eastern Province", 
           "Ha'il", "Jazan", "Madinah", "Makkah", "Najran", 
           "Northern Borders", "Riyadh", "Tabuk"),
  RR = c(3.6331406, 0.6800362, 0.6649698, 1.3338993, 0.9059756, 
         0.6910808, 1.1216577, 0.5023561, 0.9971489, 1.4107886, 
         0.7179244, 1.0816824, 0.6920513),
  EP = c(0.999999574, 0.024306301, 0.002595904, 0.940361026, 0.020938260, 
         0.006844526, 0.519099853, 0.000000000, 0.078707220, 0.845156346, 
         0.037178004, 0.389382487, 0.015469660),
  structured_effects = c(0.78884932, -0.17246398, -0.17746237, 0.35019929, -0.46516215, 
                         -0.24692756, 0.07125528, -0.08624191, 0.32013467, 0.08125823, 
                         -0.28218965, -0.08500920, -0.09632913),
  unstructured_effects = c(5.308917e-04, 4.771288e-05, -7.299302e-05, 3.320356e-04, -6.480107e-04, 
                           -2.982546e-04, -2.942489e-04, -8.681252e-05, 3.355526e-04, 2.412806e-04, 
                           -8.425624e-05, -1.060224e-04, 1.048004e-04)
)

# Get the map of Saudi Arabia using the 'rnaturalearth' package
map <- ne_states(country = "Saudi Arabia", returnclass = "sf")

# Adjust region names to match data
map$name[1] = "Eastern Province"
map$name[2] = "Northern Borders"
map$name[3] = "Al Jawf"
map$name[4] = "Najran"
map$name[5] = "Asir"
map$name[6] = "Jazan"
map$name[7] = "Tabuk"
map$name[8] = "Madinah"
map$name[9] = "Makkah"
map$name[10] = "Riyadh"
map$name[11] = "Al Qassim"
map$name[12] = "Ha'il"
map$name[13] = "Al Bahah"

# Set the correct projection
map <- st_set_crs(map, 4326)

# Merge the map data with the new data
map <- left_join(map, map_data, by = c("name" = "name"))

# Create custom color palettes for each variable
palette_rr <- colorNumeric(
  palette = c("#f2f8fa", "#b4e0e2", "#88c4cd", "#73b2b5"),
  domain = map$RR,
  na.color = "#D3D3D3"
)

palette_ep <- colorNumeric(
  palette = c("#f2f8fa", "#b4e0e2", "#88c4cd", "#73b2b5"),
  domain = map$EP,
  na.color = "#D3D3D3"
)

palette_structured_effect <- colorNumeric(
  palette = c("#f2f8fa", "#b4e0e2", "#88c4cd", "#73b2b5"),
  domain = map$structured_effect,
  na.color = "#D3D3D3"
)

palette_unstructured_effect <- colorNumeric(
  palette = c("#f2f8fa", "#b4e0e2", "#88c4cd", "#73b2b5"),
  domain = map$unstructured_effect,
  na.color = "#D3D3D3"
)

# Function to create the interactive map
# Function to create the interactive map without the legend
create_map_no_legend <- function(variable, palette, title) {
  leaflet(data = map) %>%
    addProviderTiles(providers$CartoDB.Positron) %>%
    addPolygons(
      fillColor = ~palette(variable),
      weight = 1,
      opacity = 1,
      color = "black",
      dashArray = "3",
      fillOpacity = 0.7,
      highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE),
      label = ~paste0(title, ": ", round(variable, 2))
    )
}

# Create maps for each variable without the legend
map_rr_no_legend <- create_map_no_legend(map$RR, palette_rr, "RR")
map_ep_no_legend <- create_map_no_legend(map$EP, palette_ep, "EP")
map_structured_effect_no_legend <- create_map_no_legend(map$structured_effect, palette_structured_effect, "Structured Effect")
map_unstructured_effect_no_legend <- create_map_no_legend(map$unstructured_effect, palette_unstructured_effect, "Unstructured Effect")

# Combine the maps in a grid layout with titles centered above each map
htmltools::browsable(
  tagList(
    div(
      style = "display: grid; grid-template-columns: repeat(2, 1fr); grid-template-rows: repeat(2, 1fr); gap: 20px; justify-items: center; padding: 20px; max-width: 80%; box-sizing: border-box;",
      # Title and map for RR
      div(
        style = "display: flex; flex-direction: column; align-items: center; width: 100%;",
        div(style = "color: #4f6d7a; font-size: 18px; font-weight: bold; text-align: center; margin-bottom: 10px;", "RR"),
        div(style = "width: 100%; height: 100%;", map_rr_no_legend)
      ),
      # Title and map for EP
      div(
        style = "display: flex; flex-direction: column; align-items: center; width: 100%;",
        div(style = "color: #4f6d7a; font-size: 18px; font-weight: bold; text-align: center; margin-bottom: 10px;", "EP"),
        div(style = "width: 100%; height: 100%;", map_ep_no_legend)
      ),
      # Title and map for Structured Effect
      div(
        style = "display: flex; flex-direction: column; align-items: center; width: 100%;",
        div(style = "color: #4f6d7a; font-size: 18px; font-weight: bold; text-align: center; margin-bottom: 10px;", "Structured Effect"),
        div(style = "width: 100%; height: 100%;", map_structured_effect_no_legend)
      ),
      # Title and map for Unstructured Effect
      div(
        style = "display: flex; flex-direction: column; align-items: center; width: 100%;",
        div(style = "color: #4f6d7a; font-size: 18px; font-weight: bold; text-align: center; margin-bottom: 10px;", "Unstructured Effect"),
        div(style = "width: 100%; height: 100%;", map_unstructured_effect_no_legend)
      )
    )
  )
)
```
- **Makkah, Riyadh, Eastern Province** show high relative risk, while **Najran** shows low risk.
- **Ep**: Regions like **Makkah**, **Eastern Province**, and **Riyadh** show clustering of high-risk areas.
- **Structured random effects** show minimal spatial impact, while **unstructured random effects** highlight local factors (e.g., healthcare, regional conditions).